### 2014-08-08 Docs/flowchart/flowchart.tex
start updating the flowchart -- put in the form of the equations we are currently solving.  Need to go through it to make sure it clearly shows the alternate energy formulation stuff.

>>>>

\subsection{Base State}

The stratified atmosphere is characterized by a one-dimensional
time-dependent base state, defined by a base state density, $\rho_0$,
and a base state pressure, $p_0$, in hydrostatic equilibrum:
\begin{equation}
\nabla p_0 = -\rho_0 |g| \er
\end{equation}

For the time-dependence, we will define a base state velocity, $w_0$,
which will adjust the base state from one hydrostatic equilibrum to
another in response to heating.

For convenience, we define a base state enthalphy, $h_0$, as needed
by laterally averaging the full enthalpy, $h$.

\subsection{Continuity}

Conservation of mass gives the same continuity equation we have with
compressible flow:
\begin{equation}
\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) = 0
\end{equation}

Additionally, we carry species around which can react.  The creation and destruction
of the species is described by their create rate, $\omegadot_k$, and the species
are defined by their mass fractions, $X_k \equiv \rho_k / \rho$, giving
\begin{equation}
\frac{\partial \rho X_k}{\partial t} + \nabla \cdot (\rho \Ub) = \rho \omegadot_k
\end{equation}
and
\begin{equation}
\sum_k X_k = 1
\end{equation}

\subsection{Constraint}

The equation of state is cast into an elliptic constraint on the velocity field 
by differentiating $p_0(\rho, s, X_k)$ along particle paths, giving:
\begin{equation}
\nabla \cdot (\beta_0 \Ub) = \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
\end{equation}
where
 $\beta_0$ is a density-like variable that carries background stratification, defined as
\begin{equation}
\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
\end{equation}
and 
\begin{equation}
S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext
\end{equation}
where $p_{X_k} \equiv \left. \partial p / \partial X_k
\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h / 
\partial X_k \right |_{p,T,X_{j,j\ne k}}, 
p_\rho \equiv \left. \partial p/\partial \rho \right |_{T, X_k}$, and
$\sigma \equiv p_T/(\rho c_p p_\rho)$, with
$p_T \equiv \left. \partial p / \partial T \right|_{\rho, X_k}$ and
$c_p \equiv \left.  \partial h / \partial T
\right|_{p,X_k}$ is the specific heat at constant pressure.

\subsection{Momentum}

The compressible momentum equation (written in terms of velocity is):
\begin{equation}
\rho \frac{\partial \Ub}{\partial t} + \rho \Ub \cdot \nabla \Ub + \nabla p = -\rho |g| \er
\end{equation}
Subtracting off the base state, and defining the perturbational
pressure (sometimes called the dynamic pressure) as $\pi = p -                                    
p_0$, and perturbational density as $\rho' = \rho - \rho_0$, we have:
\begin{equation}
\rho \frac{\partial \Ub}{\partial t} + \rho \Ub \cdot \nabla \Ub + \nabla \pi = -\rho' |g| \er
\end{equation}
or
\begin{equation}
\frac{\partial \Ub}{\partial t} + \Ub \cdot \nabla \Ub + \frac{1}{\rho} \nabla \pi =
   -\frac{\rho^\prime}{\rho} |g| \er
\end{equation}
This is the form of the momentum equation that we solved in papers I--III and in the multilevel
paper.

Several authors \cite{KP:2012,VLBWZ:2013} explored the idea of energy
conservation in a low Mach number system and found that an additional term (which can
look like a buoyancy) is needed in the low Mach number formulation, yielding:
\begin{equation}
\frac{\partial \Ub}{\partial t} + \Ub \cdot \nabla \Ub +
   \frac{\beta_0}{\rho} \nabla \left (\frac{p^\prime}{\beta_0} \right ) =
   -\frac{\rho^\prime}{\rho} |g| \er
\end{equation}

We decompose the full velocity field into a base
state velocity, $w_0$, that governs the base state dynamics, and a
local velocity, $\Ubt$, that governs the local dynamics, i.e.,
\begin{equation}
\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
\end{equation}
with
$\overline{(\Ubt\cdot\eb_r)} = 0$ and
$w_0 = \overline{(\Ub\cdot\eb_r)}$.
The velocity evolution equations are then
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
  r} - \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r},\label{eq:w0                                                                                                                                        
  evolution}\\ \frac{\partial\Ubt}{\partial t} &=& -\Ub\cdot\nabla\Ubt
- \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
\frac{1}{\rho}\nabla\pi +
\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r -
\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
\end{eqnarray}
where $\pi_0$ is the base state component of the perturbational pressure.
By laterally averaging to equation (\ref{eq:U divergence}),
we obtain a divergence constraint for $w_0$:
\begin{equation}
\nabla\cdot(\beta_0 w_0 \eb_r) = \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
\end{equation}
The divergence constraint for $\Ubt$ can be found by subtracting (\ref{eq:w0 divergence})
into (\ref{eq:U divergence}), resulting in
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
\end{equation}

\subsection{Energy}

Finally, we add an equation for specific enthalpy evolution to our
system.  Strictly speaking this is not necessary to close the system,
but it becomes convenient at times to define the temperature.  
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,\label{eq:enthalpy}
\end{equation}

\subsection{Base State and Perturbation Form Equations}
